---
title: "taxonomy analysis"
author: "Jigyasa_Arora"
date: "2/27/2021"
---


```{r}
library(dplyr)
library(ape)

```

## Applying the filters on marker-gene dataset (same as functional data)-
```{r}
taxonomy<-read.csv("taxonomy/all-cogs-taxonomy-contiginfo-tpm.csv",row.names = 1)
#filter1-
taxonomy_filters<-taxonomy%>%dplyr::filter(TPM>=1 & contig_length >=1000)
#filter2-
taxonomy_filters$domain<-gsub("_p__.*$","",taxonomy_filters$V3)
taxonomy_filters2<-taxonomy_filters%>%dplyr::filter(domain %in% c("d__Bacteria","d__Archaea"))
write.csv(taxonomy_filters2,file="taxonomy/all-cogs-taxonomy-contiginfo-tpm-filtered.csv")
```

## taxonomic analysis by marker gene data- Do all marker genes present on the same contig have similar taxonomic annotation + remove the mismatches
```{r}
getwd()
taxonomy_ver2<-read.csv("taxonomy/all-cogs-taxonomy-contiginfo-tpm-filtered.csv",row.names = 1) #markers present in contigs >=1000 bps and >=1TPM, similar to functional data.
colnames(taxonomy_ver2)<-c("fullproteinnames","annotation","taxonomy","contig_names","samples","gene_names","full_contig_names","contig_length","TPM","NumReads","domain")
taxonomy_ver2$fullnames<-paste(taxonomy_ver2$samples,taxonomy_ver2$contig_names,sep="_")

##do we have duplicates in COG protein annotations? 
taxonomy_ver2_proteins<-taxonomy_ver2%>%select(fullproteinnames)%>%unique()
nrow(taxonomy_ver2)
nrow(taxonomy_ver2_proteins)

print(paste0("There are no duplicates in the data as no. of proteins is same in both files ",nrow(taxonomy_ver2) , " = ", nrow(taxonomy_ver2_proteins)))

##do markers present on the same contig have similar taxonomic annotation?
taxonomy_contigs<-taxonomy_ver2%>%select(fullnames)%>%group_by(fullnames)%>%count()

print(paste0(taxonomy_contigs%>%filter(n>1)%>%nrow()/nrow(taxonomy_contigs)*100, "% of markers are present on the same contig"))



print(paste0(taxonomy_contigs%>%filter(n==2)%>%nrow()/taxonomy_contigs%>%filter(n>1)%>%nrow()*100, "% of 28.20% have two markers per contigs"))
print(paste0(taxonomy_contigs%>%filter(n==3)%>%nrow()/taxonomy_contigs%>%filter(n>1)%>%nrow()*100, "% of 28.20% have three markers per contigs"))
print(paste0(taxonomy_contigs%>%filter(n==4)%>%nrow()/taxonomy_contigs%>%filter(n>1)%>%nrow()*100, "% of 28.20% have four markers per contigs"))
print(paste0(taxonomy_contigs%>%filter(n>=5)%>%nrow()/taxonomy_contigs%>%filter(n>1)%>%nrow()*100, "% of 28.20% have five or more markers per contigs"))

#----------------------------------------------------------------------------------------------------------
##Example-for contigs with 2 marker genes-<dont have to run this step, its for standardization>
taxonomy_contigs2<-taxonomy_contigs%>%filter(n==2)
taxonomy_ver2_contigtaxa<-taxonomy_ver2%>%select(fullnames,taxonomy)%>%filter(fullnames %in% taxonomy_contigs2$fullnames) #extract markers present on same contigs

taxonomy_ver2_contigtaxa$phyla<-gsub("_c__.*$","",taxonomy_ver2_contigtaxa$taxonomy)

taxonomy_ver2_contigtaxa$class<-gsub("_o__.*$","",taxonomy_ver2_contigtaxa$taxonomy)

taxonomy_ver2_contigtaxa$order<-gsub("_f__.*$","",taxonomy_ver2_contigtaxa$taxonomy)

taxonomy_ver2_contigtaxa$family<-gsub("_g__.*$","",taxonomy_ver2_contigtaxa$taxonomy)


library(data.table)
setDT(taxonomy_ver2_contigtaxa)
taxonomy_contigs_family<-as.data.table(taxonomy_ver2_contigtaxa)[, toString(family), by = list(fullnames)]
taxonomy_contigs_family2<-setDT(taxonomy_contigs_family)[, paste0("family", 1:2) := tstrsplit(V1, ",")]


head(taxonomy_contigs_family2,5) 

taxonomy_contigs_family2$family1<-gsub(" ","",taxonomy_contigs_family2$family1)
taxonomy_contigs_family2$family2<-gsub(" ","",taxonomy_contigs_family2$family2)

taxonomy_contigs_family2<-taxonomy_contigs_family2%>%mutate(sametaxonomy=ifelse(family1==family2,"TRUE","FALSE"))
taxonomy_contigs_family2%>%filter(sametaxonomy=="TRUE")%>%nrow() 

print(paste0( taxonomy_contigs_family2%>%filter(sametaxonomy=="TRUE")%>%nrow()/nrow(taxonomy_contigs_family2)*100, "% of two markers have identical taxonomic annotation at family level when they are present on the same contig."))

#do the non-identical ones match at phyla level?
taxonomy_contigs_family3<-taxonomy_contigs_family2%>%filter(sametaxonomy=="FALSE")
taxonomy_contigs_family3$family1_phyla<-gsub("_c__.*$","",taxonomy_contigs_family3$family1)
taxonomy_contigs_family3$family2_phyla<-gsub("_c__.*$","",taxonomy_contigs_family3$family2)

taxonomy_contigs_family3<-taxonomy_contigs_family3%>%mutate(samephyla=ifelse(family1_phyla==family2_phyla,"TRUE","FALSE"))
taxonomy_contigs_family3%>%filter(samephyla=="FALSE")%>%nrow() 

#removing the marker-genes that are present on the same contig yet have different taxonomic annotation-

contigstoextract<-taxonomy_contigs_family3%>%filter(samephyla=="FALSE")%>%select(fullnames)

print(paste0( taxonomy_ver2%>%filter(fullnames %in% contigstoextract$fullnames)%>%nrow()/nrow(taxonomy_ver2)*100, "% of markers present on the same contigs donot match at phyla level taxonomic annotation."))


#------------------------------------------------------------------------------------------
##Examine all contigs with more than 2 markers-

match_columns<-function(columnnumber){
taxonomy_contigs2<-taxonomy_contigs%>%filter(n==columnnumber)
taxonomy_ver2_contigtaxa<-taxonomy_ver2%>%select(fullnames,taxonomy)%>%filter(fullnames %in% taxonomy_contigs2$fullnames) #extract markers present on same contigs

taxonomy_ver2_contigtaxa$phyla<-gsub("_c__.*$","",taxonomy_ver2_contigtaxa$taxonomy)

taxonomy_ver2_contigtaxa$class<-gsub("_o__.*$","",taxonomy_ver2_contigtaxa$taxonomy)

taxonomy_ver2_contigtaxa$order<-gsub("_f__.*$","",taxonomy_ver2_contigtaxa$taxonomy)

taxonomy_ver2_contigtaxa$family<-gsub("_g__.*$","",taxonomy_ver2_contigtaxa$taxonomy)


library(data.table)
setDT(taxonomy_ver2_contigtaxa)
taxonomy_contigs_phyla<-as.data.table(taxonomy_ver2_contigtaxa)[, toString(phyla), by = list(fullnames)]
taxonomy_contigs_phyla2<-setDT(taxonomy_contigs_phyla)[, paste0("phyla", 1:columnnumber) := tstrsplit(V1, ",")]

taxonomy_contigs_phyla2<-as.data.frame(apply(taxonomy_contigs_phyla2,2,function(x)gsub(' ', '',x))) #remove spaces in each cell.
rownames(taxonomy_contigs_phyla2)<-taxonomy_contigs_phyla2$fullnames

taxonomy_contigs_phyla2.2<-taxonomy_contigs_phyla2%>%select(-c(fullnames,V1))
test<-transform(taxonomy_contigs_phyla2.2, same = apply(taxonomy_contigs_phyla2.2, 1, function(x) length(unique(x)) == 1))

falsecolumns<-test%>%filter(same=="FALSE")

print(paste0( test%>%filter(same=="TRUE")%>%nrow()/nrow(taxonomy_contigs_phyla2)*100, "% of", columnnumber, "markers have identical taxonomic annotation at phyla level when they are present on the same contig."))
return(falsecolumns)
}

check1<-match_columns(2)
check2<-match_columns(3)
check3<-match_columns(4)
match_columns(5)
match_columns(6)
check6<-match_columns(7)
match_columns(8)
match_columns(9)
match_columns(10)
match_columns(11)
match_columns(12)
match_columns(13)
match_columns(14)
match_columns(15)
match_columns(16)
match_columns(17)
match_columns(18)
match_columns(19)
match_columns(20)
match_columns(21)
match_columns(22)
match_columns(23)
match_columns(24)
match_columns(25)
match_columns(26)
#match_columns(27)
#match_columns(28)
#match_columns(29)
#match_columns(30)
#match_columns(31)
#match_columns(32)
#match_columns(32)
#match_columns(33)
#match_columns(34)
#match_columns(35)
#match_columns(36)
#match_columns(37)
#match_columns(38)
#match_columns(39)
match_columns(40)


#---------------------------------------------------------------------------------------------
#removing the mismatched marker-genes-

taxonomy_ver3<-taxonomy_ver2%>%filter(!fullnames %in% rownames(check1))
taxonomy_ver4<-taxonomy_ver3%>%filter(!fullnames %in% rownames(check2))
taxonomy_ver5<-taxonomy_ver4%>%filter(!fullnames %in% rownames(check3))
taxonomy_ver6<-taxonomy_ver5%>%filter(!fullnames %in% rownames(check6))

write.csv(taxonomy_ver6,file="taxonomy/all-cogs-taxonomy-contiginfo-tpm-filtered-corrected.csv")
```




## get marker gene taxonomy present in 129 samples-
```{r}

#get family level taxonomy of each marker gene-
taxonomy_ver2<-read.csv("taxonomy/all-cogs-taxonomy-contiginfo-tpm-filtered-corrected.csv",row.names=1) #markers present in contigs >1000 bps and >1TPM, similar to functional data + filtered for mismatched marker-gene taxonomy.

taxonomy_ver2$family<-gsub("_g__.*$","",taxonomy_ver2$taxonomy)

#spread the data by samples and family-level taxonomy-
library(tidyr)
library(data.table)

setDT(taxonomy_ver2)
taxonomy_ver2_markertaxa<-taxonomy_ver2[, .(prokTPM = sum(TPM)), by = .(samples,family)]

taxonomy_ver2_spread<-as.data.frame(spread(taxonomy_ver2_markertaxa,family,prokTPM))
rownames(taxonomy_ver2_spread)<-taxonomy_ver2_spread$samples

#these 129 samples contain 10,000 contigs above 1000 bps, with more than 100 prokaryotic contigs.

newtree_129samples<-read.tree("1000bps_taxonomy_129taxa_tree.nwk")
dnew2<-data.frame(label=newtree_129samples$tip.label)

dnew2$label<-as.character(dnew2$label)
dnew2$runnumber<-unlist(lapply(strsplit(dnew2$label,split="-"),"[",1))
dnew2$runnumber<-gsub("_","-",dnew2$runnumber)


taxonomy_ver2_spread2<-merge(taxonomy_ver2_spread,dnew2,by.x="samples",by.y="runnumber",all.y=TRUE)
rownames(taxonomy_ver2_spread2)<-taxonomy_ver2_spread2$label


#order the rownames by termite tree to get samples with no data at this step!
taxonomy_ver2_spread2<-taxonomy_ver2_spread2[match(newtree_129samples$tip.label,rownames(taxonomy_ver2_spread2)), ]
taxonomy_ver2_spread2[is.na(taxonomy_ver2_spread2)] <-0
write.csv(taxonomy_ver2_spread2,file="metagenome-taxonomy-1000bps-bymarkersspread-rawTPM-march2021.csv")


head(taxonomy_ver2_spread2,5)
```



## data filtering for statistical analysis + clr transformation
```{r}

#clr transform the spread data-
taxonomy_ver2_spread2<-read.csv("metagenome-taxonomy-1000bps-bymarkersspread-rawTPM-march2021.csv",row.names = 1)


clr_transform<-function(filename_spread2){ 
library(propr)
filename_spread2[is.na(filename_spread2)] <-0 #convert NA to zero
#rownames(filename_spread2)<-filename_spread2$label
filename_spread2$label<-NULL
filename_spread2$samples<-NULL
#add psuedo counts to account for missing and not-present data-
filename_spread2<-filename_spread2+0.65

filename_propr<-propr(filename_spread2,metric="rho",ivar="clr")

#creating a df with all the transformed pathways and metadata-
filename_propr_df<-filename_propr@logratio

return(filename_propr_df)
}

markergene_taxonomy_129samples_clr<-clr_transform(taxonomy_ver2_spread2)


write.csv(markergene_taxonomy_129samples_clr,file="metagenome-taxonomy-1000bps-bymarkersspread-clr-march2021.csv")

##check-
head(round(sum(markergene_taxonomy_129samples_clr[1,])),3) #the overall sum per sample (i.e row) should be close to zero after clr transformation.

nrow(markergene_taxonomy_129samples_clr)
ncol(markergene_taxonomy_129samples_clr)

##filtering the taxonomy file to find atleast 3 microbes in >2.5% of termite samples. 3 taxa cutoff is based on https://bioconductor.org/packages/release/bioc/vignettes/philr/inst/doc/philr-intro.html. While 2.5% of samples is an arbitrary cut-off as different metagenome analysis pipeline has different cut-offs. Using Prevalance plots, I find that we can remove singletons, rare-taxa that might be sequencing errors, and  make sure that we can capture taxa abundant in most of the termite lineage (for example, Apicotermitinae, soil-feeding Termitidae are not sequenced too deeply). 



#from clr transformed data->re-select columns present in 5% of samples (6.5/129)-
#(previous version:2.5% of samples (3/129). Based on "basicstats_march2021.Rmd" "prevalanceThreshold" calculations.)

markergene_taxonomy_129samples_clr<-read.csv("metagenome-taxonomy-1000bps-bymarkersspread-clr-march2021.csv",row.names=1)

above5perc_function<-function(filename){
filename$label<-NULL  
filename$samplename<-NULL  

filename2<-filename #for filtering 
filename2[filename2 > 0] <- 1 
filename2[filename2 <= 0] <-0
filename_10perc<-filename2[colSums(filename2) >= 6.5]  
columns<-colnames(filename_10perc) 

filename3<-filename%>%dplyr::select(columns)
return(filename3)
}


#markergene_taxonomy_129samples_clr_2.5perc<-above2.5perc_function(markergene_taxonomy_129samples_clr)
markergene_taxonomy_129samples_clr_5perc<-above5perc_function(markergene_taxonomy_129samples_clr)

#from clr transformed data-> re-select rows with more than 3 functions 
morethan3taxa_functions<-function(filename){
 
filename$label<-NULL 
filename$samplename<-NULL
filename2<-filename
filename2[filename2 > 0] <- 1
filename2[filename2 <= 0] <-0
filename_threeabove<-filename2[rowSums(filename2[,])>=3, ]
rows<-rownames(filename_threeabove)

filename3<-subset(filename, rownames(filename) %in% rows)

return(filename3)
}

#markergene_taxonomy_129samples_clr_2.5perc_3above<-morethan3taxa_functions(markergene_taxonomy_129samples_clr_2.5perc)
markergene_taxonomy_129samples_clr_5perc_3above<-morethan3taxa_functions(markergene_taxonomy_129samples_clr_5perc)

#write.csv(markergene_taxonomy_129samples_clr_2.5perc_3above,file="taxonomy_129samples_184taxa_march2021.csv")
write.csv(markergene_taxonomy_129samples_clr_5perc_3above,file="taxonomy_129samples_123taxa_march2021.csv")
#-------------------------------------------------

```


## taxonomy stats-Moran's I and Phyl.anova-
```{r}

tree<-read.tree("1000bps_taxonomy_129taxa_tree.nwk")

functions<-read.csv("taxonomy_129samples_123taxa_march2021.csv",row.names = 1,header=TRUE)
#functions$X<-NULL
functions<-functions[match(tree$tip.label,rownames(functions)), ] #match with the tree

library(adephylo)
library(ape) 
library(phylobase) 
library(phylosignal) 
 

functions[is.na(functions)] <-0 #convert any NA to zero
functions<-functions[match(tree$tip.label,rownames(functions)), ]
p4d <- phylo4d(tree,functions)
p2<-phyloSignal(p4d = p4d, method = "I",reps=9999)

#adjusted p-value using  fdr method-
mod2 = as.data.frame(p2$pvalue)
mod2$padjusted_lambda<-p.adjust(mod2$I, method='fdr')
mod2$functions<-rownames(mod2)
mod2<-mod2%>%dplyr::mutate(significance_padjusted=ifelse(padjusted_lambda<0.05,"*",""))
mod2<-mod2%>%dplyr::mutate(significance_moranI=ifelse(I<0.05,"*",""))

write.csv(mod2,file="phylogenetic_autocorrelation_taxonomy_123columns_129samples_march2021.csv")

#-----------------------------------------------------------------------------
#Phylogenetic anova

#individually-  running the function in a list of columns-
tree<-read.tree("1000bps_taxonomy_129taxa_tree.nwk")
functions<-read.csv("taxonomy_129samples_123taxa_march2021.csv",row.names = 1,header=TRUE)
#functions$X<-NULL
functions<-functions[match(tree$tip.label,rownames(functions)), ] #match the tree
functions <-functions+100 #add a constant to every value

sampleinfo<-read.csv("taxonomy/sample_names_ids_withotherfactors_tomedit.2.csv")

library(geomorph)
phyl.anova.func2_list<-function(filename){
filename<-as.data.frame(t(filename)) #convert row to column
filename$label<-rownames(filename)
geo_functions<-merge(filename,sampleinfo,by.x="label",by.y="termite_tree_labels")
geo_functions$diet_type_number4<-as.factor(geo_functions$diet_type_number4)
geo_functions[,2]<-as.numeric(geo_functions[,2])

gdf <- geomorph.data.frame(abundance = geo_functions[,2], diet = as.factor(geo_functions$diet_type_number4), phy = tree) #create a dataframe that contains the data and phylogeny. [according to google forum]

#group_name <- names(filename)[2] #get the correct column name. column 2 is the cazyme.

geomorph.lm<-procD.lm(abundance ~ diet, iter=9999,RRPP=TRUE,effect.type = "F", SS.type = "II" ,data=gdf) #SS.type="II" is used if the groups (i.e. diet/lineage) are of uneven sizes.
mod<-anova(geomorph.lm) #to examine if there is a phylogenetic correlation between "y" and "x"
mod2<-as.data.frame(mod$table)
mod2<-mod2[1,7]
# d is distance between vector lengths is the difference in the amount of shape change per unit of size. If d is 0.05 i.e., | 0.20 - 0.15 |.  Why do this?  Two vectors can be correlated in direction - the way shape changes with size - but one group can exhibit more shape change per unit size than the other
#D tells us the effect size of the standard deviation between two groups. Positive and above 0.2-0.3 means larger positive difference, while negative and below -0.2 to -0.3 means larger negative difference.

pw<-pairwise(geomorph.lm,groups = as.factor(gdf$diet))
pw_summary<-summary(pw,confidence=0.95,show.vectors = TRUE)
pw_summary2<-as.data.frame(pw_summary$summary.table)

pw_summary2$padjust<-p.adjust(pw_summary2$`Pr > d`,method="fdr") #fdr correction of p.value
pw_summary2$diets<-rownames(pw_summary2)

#convert the output to correct formats-
names<-colnames(geo_functions[2])
pw_summary2$genename<-rep(names,times=nrow(pw_summary2))
library(tidyr)
pw_summary3<-pw_summary2%>%dplyr::select("padjust","diets","genename")
pw_spread<-pw_summary3%>%spread(diets,padjust)
pw_spread2<-cbind(pw_spread,mod2)

return(pw_spread2)
}

functions_t<-as.data.frame(t(functions))
xy.list <-  split(functions_t, seq(nrow(functions_t))) #convert df to a list

taxonomy_geomorph<-lapply(xy.list,phyl.anova.func2_list) #apply on all columns 
 
taxonomy_geomorph_df<-do.call(rbind.data.frame,taxonomy_geomorph)
 

#find p.adjust for anova-
taxonomy_geomorph_df$padjust<-p.adjust(taxonomy_geomorph_df$mod2,method="fdr")
taxonomy_geomorph_df<-taxonomy_geomorph_df%>%mutate(significant_mod2=ifelse(mod2 <=0.05,"*",""))
taxonomy_geomorph_df<-taxonomy_geomorph_df%>%mutate(significant_anova=ifelse(padjust <=0.05,"*",""))

write.csv(taxonomy_geomorph_df,file="Phylogenetic_anova_withdiet_taxonomy_123columns_129samples_march2021.csv")


```


##basic stats for taxonomy-
```{r}
#how many total bacterial lineages present-
morani<-read.csv("phylogenetic_autocorrelation_taxonomy_184columns_129samples_march2021.csv")
morani$domain<-gsub("_p__.*$","",morani$X)
morani_bacterial<-morani%>%filter(domain=="d__Bacteria")

print(paste0("There are ", nrow(morani_bacterial), " bacterial lineages in our Phylogenetic autocorrelation analysis."))

#How many phyla are present in the stats?
familylevel<-read.csv("family_levelclassifications_markergenes.csv") #generated from Moran's I dataset. All the lineages classified at family level. Those at higher level are not included here.
familylevel$phyla<-gsub("_c__.*$","",familylevel$family_level_classification)
familylevel$domain<-gsub("_p__.*$","",familylevel$phyla)

familylevel%>%select(phyla,domain)%>%group_by(phyla,domain)%>%count()
familylevel%>%filter(domain=="d__Bacteria")%>%select(phyla)%>%unique()%>%count()

print(paste0("There are ", familylevel%>%filter(domain=="d__Bacteria")%>%nrow(), "families in ",familylevel%>%filter(domain=="d__Bacteria")%>%select(phyla)%>%unique()%>%count(), "  bacterial phyla in marker-gene dataset."))


#get the Moran's I and Phylogenetic anova for the microbes classified at family level-

morani<-read.csv("phylogenetic_autocorrelation_taxonomy_184columns_129samples_march2021.csv")
morani2<-morani%>%filter(X %in% familylevel$family_level_classification)
write.csv(morani2,file="phylogenetic_autocorrelation_taxonomy_99family_129samples_march2021.csv")

anova<-read.csv("Phylogenetic_anova_withdiet_taxonomy_184columns_129samples_march2021.csv")
anova2<-anova%>%filter(genename %in% familylevel$family_level_classification)
write.csv(anova2,file="Phylogenetic_anova_withdiet_taxonomy_99family_129samples_march2021.csv")
#----------------------------------------------
#Distribution of Archaea in all samples + Mimeutermes sample-
taxonomy_clr<-read.csv("taxonomy_129samples_184taxa_march2021.csv",row.names = 1)
taxonomy_rawtpm<-read.csv("metagenome-taxonomy-1000bps-bymarkersspread-rawTPM-march2021.csv",row.names=1)

##how many archaeal families present in the rawTPM data in 124 samples-
taxonomy_rawtpm$label<-NULL
taxonomy_rawtpm$samples<-NULL
taxonomy_rawtpm[taxonomy_rawtpm > 0] <- 1
taxonomy_rawtpm[taxonomy_rawtpm <= 0] <-0
taxonomy_rawtpm_morethanone<-taxonomy_rawtpm[rowSums(taxonomy_rawtpm[,])>1, ]

taxonomy_rawtpm_cols<-colnames(taxonomy_rawtpm_morethanone)%>%as.data.frame()
colnames(taxonomy_rawtpm_cols)<-c("taxa")
taxonomy_rawtpm_cols$domain<-gsub("_p__.*$","",taxonomy_rawtpm_cols$taxa)
taxonomy_rawtpm_cols%>%filter(domain=="d__Archaea")
archaeal_columns<-taxonomy_rawtpm_cols%>%filter(domain=="d__Archaea")

taxonomy_rawtpm<-read.csv("metagenome-taxonomy-1000bps-bymarkersspread-rawTPM-march2021.csv",row.names=1)
taxonomy_raw_tpm_archaea<-taxonomy_rawtpm%>%select(as.vector(archaeal_columns$taxa))
write.csv(taxonomy_raw_tpm_archaea,file="taxonomy_rawtpm_archaealcolumns.csv")

library(stringr)
print(paste0("There are ",taxonomy_rawtpm_cols%>%filter(str_detect(taxa, "_f__") & domain=="d__Archaea")%>%nrow(), " archaeal families in our metagenomes, which are-"))
taxonomy_rawtpm_cols%>%filter(str_detect(taxa, "_f__") & domain=="d__Archaea")

##how many archaeal families present in statistical data of 129 samples-
taxonomy_clr_cols<-colnames(taxonomy_clr)%>%as.data.frame()
colnames(taxonomy_clr_cols)<-c("taxa")
taxonomy_clr_cols$domain<-gsub("_p__.*$","",taxonomy_clr_cols$taxa)
taxonomy_clr_cols%>%filter(domain=="d__Archaea")

print(paste0("There are ",taxonomy_clr_cols%>%filter(str_detect(taxa, "_f__") & domain=="d__Archaea")%>%nrow(), " archaeal families in 2.5% of termite samples, which are-"))
taxonomy_clr_cols%>%filter(str_detect(taxa, "_f__") & domain=="d__Archaea")
#--------------------------------------------------------------------------------
##Overall relative abundance of Archaea-
taxonomy_rawtpm<-read.csv("metagenome-taxonomy-1000bps-bymarkersspread-rawTPM-march2021.csv",row.names=1)
taxonomy_rawtpm$samples<-NULL
taxonomy_long<-gather(taxonomy_rawtpm, taxa, tpm, -(label), factor_key=TRUE)%>%as.data.frame()
taxonomy_long$domain<-gsub("_p__.*$","",taxonomy_long$taxa)
taxonomy_long<-taxonomy_long%>%filter(tpm>0)

library(data.table)
setDT(taxonomy_long)
taxonomy_long_domains<-taxonomy_long[, .(tpm = sum(tpm)), by = .(domain)]

taxonomy_long_domains[2,2]/sum(taxonomy_long_domains$tpm)*100

print(paste0("Proportion of Archaea in 124 metagenomes are ", taxonomy_long_domains[1,2]/sum(taxonomy_long_domains$tpm)*100 ,"%"))

#-------------------------------------------------------------------------------
##Which sample has the most relative abundance of archaea?

library(data.table)
setDT(taxonomy_long)
taxonomy_long_sampledomains<-taxonomy_long[, .(tpm = sum(tpm)), by = .(label,domain)]
taxonomy_long_sampledomains_archaea<-taxonomy_long_sampledomains%>%filter(domain=="d__Archaea")
max(taxonomy_long_sampledomains_archaea$tpm)
taxonomy_long_sampledomains_archaea%>%filter(tpm==max(taxonomy_long_sampledomains_archaea$tpm)) #extract the samplename with most archaeal TPM in it.

write.csv(taxonomy_long_sampledomains,file="taxonomy_rawtpm_domainlevel.csv")

mimeutermes<-taxonomy_long%>%filter(label %in% c("272_32-CIVT017-Nasutitermitinae-Mimeutermes_sorex"))
mimeutermes_bathy<-mimeutermes%>%filter(str_detect(taxa, "Bathyarchaeia"))

print(paste0("Proportion of Bathyarchaeia in Mimeutermes sample is ", sum(mimeutermes_bathy$tpm)/sum(mimeutermes$tpm)*100))

#NOTE-Mimeutermes has the most no. of Archaeal relative abundance in all samples.
#---------------------------------------------------------------------------------------
##do all soil-feeders have higher Archaeal relative abundance than other diets?
sampleinfo<-read.csv("taxonomy/sample_names_ids_withotherfactors_tomedit.2.csv")
taxonomy_long_diet<-merge(taxonomy_long,sampleinfo,by.x="label",by.y="termite_tree_labels")

library(data.table)
setDT(taxonomy_long_diet)
taxonomy_long_dietdomains<-taxonomy_long_diet[, .(tpm = sum(tpm)), by = .(diet_type_number4,domain)]
taxonomy_long_dietdomains<-taxonomy_long_dietdomains%>%group_by(diet_type_number4)%>%mutate(overallsum=sum(tpm))
taxonomy_long_dietdomains$perc<-taxonomy_long_dietdomains$tpm/taxonomy_long_dietdomains$overallsum*100
write.csv(taxonomy_long_dietdomains,file="taxonomy_rawtpmsummed_perdiet.csv")
taxonomy_long_dietdomains%>%filter(domain=="d__Archaea")


#NOTE-Archaea are more abundant in soil-feeding termites, followed by fungal-cultivating termites.


#archaeal distribution for all higher termites together-
library(data.table)
setDT(taxonomy_long_diet)
taxonomy_long_dietdomains<-taxonomy_long_diet[, .(tpm = sum(tpm)), by = .(higher_lower,domain)]
taxonomy_long_dietdomains<-taxonomy_long_dietdomains%>%group_by(higher_lower)%>%mutate(overallsum=sum(tpm))
taxonomy_long_dietdomains$perc<-taxonomy_long_dietdomains$tpm/taxonomy_long_dietdomains$overallsum*100
write.csv(taxonomy_long_dietdomains,file="taxonomy_rawtpmsummed_perhigherlower.csv")
taxonomy_long_dietdomains%>%filter(domain=="d__Archaea")


```





##16S and marker gene taxonomy comparison-
```{r}
ncbi<-read.csv("taxonomy/16S_newdata_89samples_408columns.csv") #OTUs summed at family level
markergene<-read.csv("metagenome-taxonomy-1000bps-bymarkersspread-rawTPM-march2021.csv")

#how many samples are in common between the two datasets?
ncbi_markers<-merge(ncbi,markergene,by.x="termite_samples","X") #74 termite samples in common.
rownames(ncbi_markers)<-ncbi_markers$termite_samples

#how many family-level taxonomy in common between the two datasets?
library(stringr)
taxa_16s<-colnames(ncbi)%>%as.data.frame()
colnames(taxa_16s)<-c("taxa")
taxa_16s$taxaimproved<-taxa_16s$taxa
taxa_16s$taxaimproved<-gsub("_fa$","",taxa_16s$taxaimproved)
taxa_16s$taxaimproved<-gsub("_unclassified$","",taxa_16s$taxaimproved)
taxa_16s$taxaimproved<-gsub("_uncultured$","",taxa_16s$taxaimproved)
taxa_16s$taxaimproved<-gsub("_$","",taxa_16s$taxaimproved)
taxa_16s$taxaimproved<-gsub("_soil_group$","",taxa_16s$taxaimproved)
taxa_16s$taxaimproved<-gsub("_termite_group$","",taxa_16s$taxaimproved)
taxa_16s$taxaimproved<-gsub("_uncultured_uncultured_uncultured_or$","",taxa_16s$taxaimproved)
taxa_16s$taxaimproved<-gsub("_or$","",taxa_16s$taxaimproved)
taxa_16s$taxaimproved<-gsub("_group$","",taxa_16s$taxaimproved)
taxa_16s$taxaimproved<-gsub("_uncultured$","",taxa_16s$taxaimproved)
taxa_16s$taxaimproved<-gsub("_lineage$","",taxa_16s$taxaimproved)
taxa_16s$taxaimproved<-gsub("_gut$","",taxa_16s$taxaimproved)

taxa_16s$taxonomyofinterest<-gsub("^.*_", "",taxa_16s$taxaimproved)

#are there any duplicates in "taxonomyofinterest" column for 16S-
dups<-taxa_16s%>%select(taxonomyofinterest)%>%group_by(taxonomyofinterest)%>%count()%>%filter(n>1)
taxa_16s_dups<-taxa_16s%>%filter(taxonomyofinterest %in% dups$taxonomyofinterest)
write.csv(taxa_16s_dups,file="duplicates_16s_taxonomy.csv")
#NOTE-if there are duplicates, manually add them in the "taxonomy/16S_newdata_89samples_408columns.csv" file

taxa_markers<-colnames(markergene)%>%as.data.frame()
colnames(taxa_markers)<-c("taxa")
taxa_markers$taxaimproved<-taxa_markers$taxa
taxa_markers$taxaimproved<-gsub("Treponemataceae_B$","TreponemataceaeB",taxa_markers$taxaimproved)
taxa_markers$taxaimproved<-gsub("Firmicutes_A$","FirmicutesA",taxa_markers$taxaimproved)
taxa_markers$taxaimproved<-gsub("Firmicutes_B$","FirmicutesB",taxa_markers$taxaimproved)
taxa_markers$taxaimproved<-gsub("Desulfotomaculaceae_A$","DesulfotomaculaceaeA",taxa_markers$taxaimproved)
taxa_markers$taxaimproved<-gsub("Bacillaceae_G$","BacillaceaeG",taxa_markers$taxaimproved)
taxa_markers$taxonomyofinterest<-gsub("^.*_", "",taxa_markers$taxaimproved)


commontaxa<-merge(taxa_markers,taxa_16s,by="taxonomyofinterest") #160 family-level taxa in common.
commontaxa_vector<-c(as.vector(commontaxa$taxa.x),as.vector(commontaxa$taxa.y))

names<-names(ncbi_markers)[(names(ncbi_markers) %in% commontaxa_vector)]
ncbi_markers_subset <- ncbi_markers[, names]

write.csv(ncbi_markers_subset,file="taxonomy_16S_markers_rawcounts.csv")
#NOTE- we have 74 termite samples with 145 common family-level taxa.
#-------------------------------------------------------------------------------------------
##extract out the "others" column in each file-

#ncbi 16S-
ncbi_long<-gather(ncbi,taxa,count,2:nrow(ncbi))
ncbi_long<-ncbi_long%>%mutate(presence=ifelse(taxa %in% names,"common","others"))
ncbi_long<-ncbi_long%>%filter(count!=0)

markers_long<-gather(markergene,taxa,count,3:326)
markers_long<-markers_long%>%mutate(presence=ifelse(taxa %in% names,"common","others"))
markers_long$count<-as.numeric(markers_long$count)
markers_long<-markers_long%>%filter(count!=0)

#sum by "others"
setDT(ncbi_long)
ncbi_long_others<-ncbi_long[, .(count = sum(count)), by = .(termite_samples,presence)]
ncbi_long_others<-ncbi_long_others%>%filter(presence=="others")


setDT(markers_long)
markers_long_others<-markers_long[, .(count = sum(count)), by = .(X,presence)]
markers_long_others<-markers_long_others%>%filter(presence=="others")
#-------------------------------------------------------------------------------------------
##join all the data in one-
ncbi_markers_subset<-read.csv("taxonomy_16S_markers_rawcounts.csv")
ncbi_markers_subset2<-merge(ncbi_markers_subset,ncbi_long_others,by.x="X",by.y="termite_samples")
ncbi_markers_subset3<-merge(ncbi_markers_subset2,markers_long_others,by="X")
ncbi_markers_subset3$presence.y<-NULL
ncbi_markers_subset3$presence.x<-NULL
names(ncbi_markers_subset3)[names(ncbi_markers_subset3) == "count.x"] <- "s16_others"
names(ncbi_markers_subset3)[names(ncbi_markers_subset3) == "count.y"] <- "markers_others"

write.csv(ncbi_markers_subset3,file="taxonomy_16S_markers_others_rawcounts.csv")
#-------------------------------------------------------------------------------------------
##CLR-transform the data 
ncbi_markers_subset3<-read.csv("taxonomy_16S_markers_others_rawcounts.csv",row.names=1)
rownames(ncbi_markers_subset3)<-ncbi_markers_subset3$X
ncbi_markers_subset3$X<-NULL

#perform clr-transformation on 16S and markergenes separately.
columns<-colnames(ncbi_markers_subset3)%>%as.data.frame()
colnames(columns)<-c("cols")
columns<-columns%>%mutate(columnids=ifelse(str_detect(cols, "^Bacteria|^Archaea|^s16"),"s16","markers"))
columns_16s<-columns%>%filter(columnids=="s16" | cols=="termite_samples")
columns_markers<-columns%>%filter(columnids=="markers" | cols=="termite_samples")
#markers-
library(stringr)
markers_only<-ncbi_markers_subset3[, names(ncbi_markers_subset3) %in% columns_markers$cols]
rownames(markers_only)<-markers_only$termite_samples
#16S-
ncbi_only<-ncbi_markers_subset3[, names(ncbi_markers_subset3) %in% columns_16s$cols]
rownames(ncbi_only)<-ncbi_only$termite_samples

clr_transform<-function(filename_spread2){ 
library(propr)
filename_spread2[is.na(filename_spread2)] <-0 #convert NA to zero
#rownames(filename_spread2)<-filename_spread2$label
filename_spread2$samples<-NULL
filename_spread2$termite_samples<-NULL
#add psuedo counts to account for missing and not-present data-
filename_spread2<-filename_spread2+0.65

filename_propr<-propr(filename_spread2,metric="rho",ivar="clr")
#creating a df-
filename_propr_df<-filename_propr@logratio

return(filename_propr_df)
}

ncbi_only_clr<-clr_transform(ncbi_only)
markers_only_clr<-clr_transform(markers_only)


#order the df by termite tree
tree<-read.tree("basicstats/rna12-16S_concensus_may2020_newnames.nwk")
newtree<-keep.tip(tree,as.vector(rownames(ncbi_only_clr)))

ncbi_only_clr<-ncbi_only_clr[match(newtree$tip.label,rownames(ncbi_only_clr)), ]
markers_only_clr<-markers_only_clr[match(newtree$tip.label,rownames(markers_only_clr)), ]

#cbind the data
ncbi_markers_clr<-cbind(ncbi_only_clr,markers_only_clr)%>%as.data.frame()
write.csv(ncbi_markers_clr,file="taxonomy_16S_markers_others_clr.csv") #no filters applied!
#-------------------------------------------------------------------------------------------
above5perc_function<-function(filename){
filename$label<-NULL  
filename$samplename<-NULL  

filename2<-filename #for filtering 
filename2[filename2 > 0] <- 1 
filename2[filename2 <= 0] <-0
filename_10perc<-filename2[colSums(filename2) >= 4]  #4/74=5%
columns<-colnames(filename_10perc) 

filename3<-filename%>%dplyr::select(columns)
return(filename3)
}

#ncbi_markers_clr_5perc<-above5perc_function(ncbi_markers_clr)

#from clr transformed data-> re-select rows with more than 3 functions 
morethan3taxa_functions<-function(filename){
 
filename$label<-NULL 
filename$samplename<-NULL
filename2<-filename
filename2[filename2 > 0] <- 1
filename2[filename2 <= 0] <-0
filename_threeabove<-filename2[rowSums(filename2[,])>=3, ]
rows<-rownames(filename_threeabove)

filename3<-subset(filename, rownames(filename) %in% rows)

return(filename3)
}

#ncbi_markers_clr_5perc_3above<-morethan3taxa_functions(ncbi_markers_clr_5perc)
write.csv(ncbi_markers_clr,file="taxonomy_16S_markers_others_clr.csv") #no filters applied!
#----------------------------------------------------------------------------------------------------
#reorder the columns in the dataframe-
ncbi_markers_clr<-read.csv("taxonomy_16S_markers_others_clr.csv",row.names=1)

#get the new names-
columns<-colnames(ncbi_markers_clr)%>%as.data.frame()
colnames(columns)<-c("originalcolumns")
columns$newcolumns<-columns$originalcolumns
columns$newcolumns<-gsub("_fa$","",columns$newcolumns)
columns$newcolumns<-gsub("_unclassified$","",columns$newcolumns)
columns$newcolumns<-gsub("_uncultured$","",columns$newcolumns)
columns$newcolumns<-gsub("_$","",columns$newcolumns)
columns$newcolumns<-gsub("_soil_group$","",columns$newcolumns)
columns$newcolumns<-gsub("_termite_group$","",columns$newcolumns)
columns$newcolumns<-gsub("_uncultured_uncultured_uncultured_or$","",columns$newcolumns)
columns$newcolumns<-gsub("_or$","",columns$newcolumns)
columns$newcolumns<-gsub("_group$","",columns$newcolumns)
columns$newcolumns<-gsub("_uncultured$","",columns$newcolumns)
columns$newcolumns<-gsub("_lineage$","",columns$newcolumns)
columns$newcolumns<-gsub("_gut$","",columns$newcolumns)
columns$newcolumns<-gsub("Treponemataceae_B$","TreponemataceaeB",columns$newcolumns)
columns<-columns%>%mutate(columnids=ifelse(str_detect(newcolumns, "^Bacteria|^Archaea|^s16"),"s16","markers"))
columns$newcolumns2<-gsub("^.*_", "",columns$newcolumns)
columns$finalcolumns<-paste(columns$columnids,columns$newcolumns2,sep="_")

#get the full taxonomy in the column header
columns2<-merge(columns,taxa_markers,by.x="newcolumns2",by.y="taxonomyofinterest",all.x = TRUE)
columns2$finalcolumns2<-paste(columns2$columnids,columns2$taxaimproved,sep="_")
columns2<-columns2%>%mutate(finalcolumns2=ifelse(is.na(taxa),as.character(finalcolumns),as.character(finalcolumns2)))
write.csv(columns2,file="taxonomy_columns.csv") ##NOTE-CHECK this df manually before running the next step

#rename the columns of the markergene-16s data on "columns" df-
ncbi_markers_clr2<- ncbi_markers_clr%>% rename_at(vars(columns2$originalcolumns), ~ columns2$finalcolumns2)
#order the "columns" df by taxonomy-
columns2<-columns2[order(columns2$newcolumns2), ]
#get the same order in the markergene-16S data-
ncbi_markers_clr2 <- ncbi_markers_clr2[columns2$finalcolumns2]

  
write.csv(ncbi_markers_clr2,file="taxonomy_16S_markers_others_clr_ordered.csv")

```

##16S tests-plotting the 16S data alone-top 10 phyla present in 74 samples
```{r}

ncbi<-read.csv("taxonomy/16S_newdata_finaldataset.csv",row.names=1) #OTUs from Vincent Herve


#get the OTUs summed by phyla per sample-
library(tidyr)
ncbi_long<-gather(ncbi, termite_samples, rawreads, -c(OTUs,taxonomy), factor_key=TRUE)
ncbi_long<-ncbi_long%>%filter(rawreads>0)
ncbi_long$termite_samples<-gsub("\\..*$","",ncbi_long$termite_samples)
ncbi_long$termite_samples<-gsub("^X","",ncbi_long$termite_samples)
ncbi_long$termite_samples<-gsub("_","-",ncbi_long$termite_samples)

#get 74 samples out-
samples<-read.csv("taxonomy/74samples_16s_markers.csv",header=FALSE)
ncbi_long2<-ncbi_long%>%filter(termite_samples %in% samples$V1)

#get phyla columnt out-
ncbi_long2$taxa1<-unlist(lapply(strsplit(as.character(ncbi_long2$taxonomy),split="_"),"[",1))
ncbi_long2$taxa2<-unlist(lapply(strsplit(as.character(ncbi_long2$taxonomy),split="_"),"[",2))
ncbi_long2$phyla<-paste(ncbi_long2$taxa1,ncbi_long2$taxa2,sep="_")

#sum by phyla and spread by phyla-
ncbi_long3<-ncbi_long2%>%select(termite_samples,rawreads,phyla)
setDT(ncbi_long3)
ncbi_long3sum<-ncbi_long3[, .(prokreads = sum(rawreads)), by = .(termite_samples,phyla)]

#spread by phyla and termite samples-
ncbi_long3sum_spread<-as.data.frame(spread(ncbi_long3sum,phyla,prokreads))
rownames(ncbi_long3sum_spread)<-ncbi_long3sum_spread$termite_samples

#remove columns that are zero i.e. phyla absent in all samples-
abovezero_function<-function(filename){
filename$X<-NULL 
filename$label<-NULL 
filename$termite_samples<-NULL
filename[is.na(filename)] <-0
filename2<-filename #for filtering
filename2[filename2 > 0] <- 1 
filename2[filename2 <= 0] <-0
filename_10perc<-filename2[colSums(filename2) >0] 
columns<-colnames(filename_10perc) 

filename3<-filename%>%dplyr::select(columns)
return(filename3)
}

ncbi_long3sum_spread_above0<-abovezero_function(ncbi_long3sum_spread)

##plot1- clr transformed data only-
#clr transform the data-
clr_transform<-function(filename_spread2){ 
library(propr)
filename_spread2[is.na(filename_spread2)] <-0 #convert NA to zero
#rownames(filename_spread2)<-filename_spread2$label
filename_spread2$samples<-NULL
filename_spread2$termite_samples<-NULL
#add psuedo counts to account for missing and not-present data-
filename_spread2<-filename_spread2+0.65

filename_propr<-propr(filename_spread2,metric="rho",ivar="clr")
#creating a df-
filename_propr_df<-filename_propr@logratio

return(filename_propr_df)
}

ncbi2_clr<-clr_transform(ncbi_long3sum_spread_above0)
ncbi2_clr$X<-rownames(ncbi2_clr)

#long format-
library(tidyr)

taxa_long<-gather(ncbi2_clr, phyla, clr, -c(X), factor_key=TRUE)

#add sample info-
sampleinfo<-read.csv("taxonomy/sample_names_ids_withotherfactors_tomedit.2.csv")

taxa_long_info<-merge(taxa_long,sampleinfo,by.x="X",by.y="samples")
taxa_long_info2<-taxa_long_info%>%dplyr::filter(clr>0)


#get the main phyla present-
#phylagroups<-taxa_long_info2%>%dplyr::select(phyla)%>%dplyr::group_by(phyla)%>%dplyr::count()%>%dplyr::arrange(desc(n))

setDT(taxa_long_info2)
phylagroups<-taxa_long_info2[, .(prokclr = sum(clr)), by = .(phyla)]
phylagroups<-phylagroups%>%arrange(desc(prokclr))

#keeping top 10 microbial phyla + put others in "others" columns for plotting-
notothers<-c("Bacteria_Spirochaetota","Bacteria_Firmicutes","Bacteria_Bacteroidota","Bacteria_Proteobacteria","Bacteria_Actinobacteriota","Bacteria_Fibrobacterota","Bacteria_Synergistota","Bacteria_Desulfobacterota","Bacteria_Rs-K70","Bacteria_Planctomycetota","Bacteria_Elusimicrobiota") #the top10 phyla + Elusimicrobiota to match the markergene results.

taxa_long_info2<-taxa_long_info2%>%mutate(newphyla=ifelse(phyla %in% notothers,as.character(phyla),"Others"))

taxa_long_info2$newphyla<-gsub("^.*_","",taxa_long_info2$newphyla)
taxa_long_info2$newphyla<-gsub("\\.","",taxa_long_info2$newphyla)
taxa_long_info2$newphyla<-gsub("-","",taxa_long_info2$newphyla)
#--------------------------------------
##plot1-
library(ggplot2)

#colors used for microbial taxa-based on https://github.com/katiejolly/nationalparkcolors/blob/master/R/colors.R
##BASED ON MAG TREE JAN2021-
colors<-c(Euryarchaeota="#F6955E",Acidobacteriota="#2c5aa0", Actinobacteriota="#00aa88",Bacteroidota="#7c916f", Campylobacterota="#EC8FA3", Deferribacterota="#8C9D57",Elusimicrobiota="#4E5462", Firmicutes="#502d16",Myxococcota="#B23539",Planctomycetota="#008033",Proteobacteria="#916f7c",Fibrobacterota="#2E8289",Spirochaetota="#d4aa00",Verrucomicrobiota="#548F01",Desulfobacterota="#A1045A",Halobacteriota="#A4BED5",Thermoplasmatota="#CD5733",Patescibacteria="#359F8B",Synergistota="#802729",Chloroflexi="#EEEAA0",Cloacimonadota="#8D7F99",Cyanobacteria="#163343",Campilobacterota="#A49A69",Margulisbacteria="#73652D",Firmicutes_C="#B4674E",Methanobacteriota="#FED789",Myxococcota="#40663F",RsK70="#80792B")

dietnames<-c("1"="lower termites (LT)","2"="wood-feeding higher termites (WF)","3"="soil feeding termites (SF)","5"="fungal-cultivating termites (FC)")

#overall theme-
library(ggplot2)
addline_format <- function(x,...){
    gsub('\\s','\n',x)
}

theme_set(theme_minimal() + theme(legend.position = "right", axis.title =  element_text(size = 20), axis.text =  element_text(size = 15), strip.text = element_text(size = 20)))

#order the df according to the phyla column for the stacked boxplot-
desired_order <- c("Actinobacteriota","Bacteroidota","Desulfobacterota","Elusimicrobiota","Fibrobacterota","Firmicutes","Planctomycetota","Proteobacteria","RsK70","Spirochaetota","Synergistota","Others")
# Re-order the levels
taxa_long_info2$newphyla <- factor( as.character(taxa_long_info2$newphyla), levels=desired_order )
# Re-order the data.frame
taxa_long_info2 <- taxa_long_info2[order(taxa_long_info2$newphyla),]

p<-taxa_long_info2%>%ggplot( aes(fill=newphyla, y=clr, x=as.factor(diet_type_number4))) + geom_bar(stat="identity",position="fill",alpha=1)+scale_fill_manual(values=colors,na.value="black")+scale_x_discrete(labels=addline_format(dietnames))+labs(y="Relative abundance of top 10 microbial phyla",x="Termite diet",title="16S",fill="Phyla")+theme(plot.title = element_text(hjust = 0.5))

#------------------------------------------------------------------------------------------------
#------------------------------------------------------------------------------------------------


```


##marker-gene tests-plotting the markergene data alone-top 10 phyla present in 74 samples
```{r}
markergenes<-read.csv("taxonomy/all-cogs-taxonomy-contiginfo-tpm-filtered-corrected.csv",row.names=1) #generated from "taxonomy_march2021.Rmd"
markergenes$phyla<-gsub("_c__.*$","",markergenes$taxonomy)


library(tidyr)
library(data.table)

#sum by phyla and samples
markergenes2<-markergenes%>%dplyr::select(samples,phyla,TPM)
setDT(markergenes2)
markergenes2sum<-markergenes2[, .(prokTPM = sum(TPM)), by = .(samples,phyla)]

#spread by family + extract 74 samples out-
markergenes2sum_spread<-as.data.frame(spread(markergenes2sum,phyla,prokTPM))
rownames(markergenes2sum_spread)<-markergenes2sum_spread$samples

samplenames<-read.csv("taxonomy/74samples_16s_markers.csv",header=FALSE)
markergenes2sum_spread2<-markergenes2sum_spread%>%filter(samples %in% samplenames$V1)


#remove zero values
abovezero_function<-function(filename){
filename$X<-NULL 
filename$label<-NULL
filename$samples<-NULL
filename$termite_samples<-NULL
filename[is.na(filename)] <-0
filename2<-filename #for filtering
filename2[filename2 > 0] <- 1 
filename2[filename2 <= 0] <-0
filename_10perc<-filename2[colSums(filename2) >0] 
columns<-colnames(filename_10perc) 

filename3<-filename%>%dplyr::select(columns)
return(filename3)
}

markergenes2sum_spread2_above0<-abovezero_function(markergenes2sum_spread2)

##plot1- clr transformed data only-
#clr transform the data-
clr_transform<-function(filename_spread2){ 
library(propr)
filename_spread2[is.na(filename_spread2)] <-0 #convert NA to zero
#rownames(filename_spread2)<-filename_spread2$label
filename_spread2$samples<-NULL
filename_spread2$termite_samples<-NULL
#add psuedo counts to account for missing and not-present data-
filename_spread2<-filename_spread2+0.65

filename_propr<-propr(filename_spread2,metric="rho",ivar="clr")
#creating a df-
filename_propr_df<-filename_propr@logratio

return(filename_propr_df)
}

markergene_clr<-clr_transform(markergenes2sum_spread2_above0)
markergene_clr$X<-rownames(markergene_clr)

#long format-
library(tidyr)

markertaxa_long<-gather(markergene_clr, phyla, clr, -c(X), factor_key=TRUE)

#add sample info-
sampleinfo<-read.csv("taxonomy/sample_names_ids_withotherfactors_tomedit.2.csv")

markertaxa_long_info<-merge(markertaxa_long,sampleinfo,by.x="X",by.y="samples")
markertaxa_long_info2<-markertaxa_long_info%>%dplyr::filter(clr>0)


#get the main phyla present-
#phylagroups<-taxa_long_info2%>%dplyr::select(phyla)%>%dplyr::group_by(phyla)%>%dplyr::count()%>%dplyr::arrange(desc(n))

setDT(markertaxa_long_info2)
phylagroups<-markertaxa_long_info2[, .(prokclr = sum(clr)), by = .(phyla)]
phylagroups<-phylagroups%>%arrange(desc(prokclr))

#keeping top 10 microbial phyla + put others in "others" columns for plotting-
notothers<-c("d__Bacteria_p__Spirochaetota","d__Bacteria_p__Bacteroidota","d__Bacteria_p__Proteobacteria","d__Bacteria_p__Firmicutes_A","d__Bacteria_p__Firmicutes","d__Bacteria_p__Fibrobacterota","d__Bacteria_p__Actinobacteriota","d__Bacteria_p__Desulfobacterota","d__Bacteria_p__Elusimicrobiota","d__Bacteria_p__Planctomycetota","d__Bacteria_p__Synergistota") #the top10 phyla+ Synergistota to match 16S.

markertaxa_long_info2<-markertaxa_long_info2%>%mutate(newphyla=ifelse(phyla %in% notothers,as.character(phyla),"Others"))

markertaxa_long_info2$newphyla<-gsub("^.*p__","",markertaxa_long_info2$newphyla)
markertaxa_long_info2$newphyla<-gsub("_","",markertaxa_long_info2$newphyla)


#----------------------------------------------------------------------------------------
##plot1-

library(ggplot2)

#colors used for microbial taxa-based on https://github.com/katiejolly/nationalparkcolors/blob/master/R/colors.R
##BASED ON MAG TREE JAN2021-
colors<-c(Euryarchaeota="#F6955E",Acidobacteriota="#2c5aa0", Actinobacteriota="#00aa88",Bacteroidota="#7c916f", Campylobacterota="#EC8FA3", Deferribacterota="#8C9D57",Elusimicrobiota="#4E5462", Firmicutes="#502d16",Myxococcota="#B23539",Planctomycetota="#008033",Proteobacteria="#916f7c",Fibrobacterota="#2E8289",Spirochaetota="#d4aa00",Verrucomicrobiota="#548F01",Desulfobacterota="#A1045A",Halobacteriota="#A4BED5",Thermoplasmatota="#CD5733",Patescibacteria="#359F8B",Synergistota="#802729",Chloroflexi="#EEEAA0",Cloacimonadota="#8D7F99",Cyanobacteria="#163343",Campilobacterota="#A49A69",Margulisbacteria="#73652D",FirmicutesA="#B4674E",Methanobacteriota="#FED789",Myxococcota="#40663F",RsK70="#80792B")

dietnames<-c("1"="lower termites (LT)","2"="wood-feeding higher termites (WF)","3"="soil feeding termites (SF)","5"="fungal-cultivating termites (FC)")

#overall theme-
library(ggplot2)
addline_format <- function(x,...){
    gsub('\\s','\n',x)
}

theme_set(theme_minimal() + theme(legend.position = "right", axis.title =  element_text(size = 20), axis.text =  element_text(size = 15), strip.text = element_text(size = 20)))

#order the df according to the phyla column for the stacked boxplot-
desired_order <- c("Actinobacteriota","Bacteroidota","Desulfobacterota","Elusimicrobiota","Fibrobacterota","Firmicutes","FirmicutesA","Planctomycetota","Proteobacteria","Spirochaetota","Synergistota","Others")
# Re-order the levels
markertaxa_long_info2$newphyla <- factor( as.character(markertaxa_long_info2$newphyla), levels=desired_order )
# Re-order the data.frame
markertaxa_long_info2 <- markertaxa_long_info2[order(markertaxa_long_info2$newphyla),]

q<-markertaxa_long_info2%>%ggplot( aes(fill=newphyla, y=clr, x=as.factor(diet_type_number4))) + geom_bar(stat="identity",position="fill",alpha=1)+scale_fill_manual(values=colors,na.value="black")+scale_x_discrete(labels=addline_format(dietnames))+labs(y="Relative abundance of top 10 microbial phyla",x="Termite diet",title="Marker-genes",fill="Phyla")+theme(plot.title = element_text(hjust = 0.5))

#-------------------------------------------------------------------------------------------

library(ggpubr)
figure <- ggarrange(p, q,
                    labels = c("A", "B"),
                    ncol = 2, nrow = 1,
                    common.legend = FALSE, legend = "right")

library(svglite)
ggsave(file="16S_markers_taxonomy_march2021_new.pdf",plot=figure,height=8.5,width=13,dpi=800,units="in")

```



